---
title: "Comparing sound categories"
author: "Annebelle"
date: "2023-12-12"
output: html_document
---

```{r}
soundsMW<-read.delim("C:/Users/P310512/Documents/Groningen/Sound library/HYD92_LAUW-F1.txt")
soundsJdB<-as.data.frame(May_Won)
MayJdB<-as.data.frame(readxl::read_excel("G:/Shared drives/WaddenSea fish sounds/final log data janneke/may1/total may1 file for R.xlsx"))
JdB_Mayon<-subset(MayJdB,location=="W_on")
```

#Calculate start times for Maryann's data
```{r}
sttime<-split(soundsMW$Begin.File,".")
sttime<-sapply(strsplit(as.character(soundsMW$Begin.File), "6047."), "[", 2)
sttime2<-sapply(strsplit(sttime,".wav"),"[",1)
soundsMW$filetimes<-as.POSIXct(sttime2,format="%y%m%d%H%M%S",tz="UTC")
#soundsMW$startUTC<-soundsMW$filetimes[1]+soundsMW$Begin.Time..s.
soundsMW$startUTC<-soundsMW$filetimes+(soundsMW$Beg.File.Samp..samples./24000)
#soundsMW$endUTC<-soundsMW$filetimes[1]+soundsMW$End.Time..s.
```

#Sort data frames to start times of calls
```{r}
soundsMW2<-soundsMW[order(soundsMW$startUTC),]
soundsJdB2<-soundsJdB[order(soundsJdB$`Start time`),]
soundsJdB2<-JdB_Mayon[order(JdB_Mayon$`Start time`),]
```

#Save subsetted files
```{r}
write.csv(soundsMW2,file="G:/Shared drives/WaddenSea fish sounds/LAUW_May1_W_ON_Maryann.csv")
write.csv(soundsJdB2,file="G:/Shared drives/WaddenSea fish sounds/LAUW_May1_W_ON_Janneke.csv")

```

#Match start times
```{r}
b<-match(soundsMW2$startUTC,soundsJdB2$`Start time`)
na.omit(b)
require(ggplot2)

soundsMW2$y<-1
soundsJdB2$y<-0.9
ggplot(data=subset(soundsMW2,soundcode=="BuBz"),aes(x=startUTC,y=y))+geom_point()+geom_point(data=subset(soundsJdB2),aes(x=`Start time`,y=y),colour="red")+ylim(0.5,1.5)
  xlim(as.POSIXct(c("2023-05-08 00:00:00","2023-05-08 12:25:59"),format="%Y-%m-%d %H:%M:%S",tz="UTC"))
```

```{r}
bgJdB<-subset(soundsJdB2,Comments=="bubble growl")
bgMW<-subset(soundsMW2,soundcode=="Sea-Gr")

sfgJdB<-subset(soundsJdB2,Comments=="striped fish grunt" | Comments=="solid fish grunt")
sfgMW<-subset(soundsMW2,soundcode=="Gr")
ggplot(data=sfgMW,aes(x=startUTC,y=y))+geom_point()+geom_point(data=sfgJdB,aes(x=`Start time`,y=y),colour="red")+ylim(0.5,1.5)
  xlim(as.POSIXct(c("2023-05-04 12:00:00","2023-05-06 00:25:59"),format="%Y-%m-%d %H:%M:%S",tz="UTC"))
```

```{r}
head(soundsMW2$startUTC)
head(soundsJdB2$`Start time`)
```



Finally aligned the sound categories through the sound library. Now I will create new log files with Maryann's naming scheme.
```{r}
# Read in the log file
logfile<-as.data.frame(readxl::read_excel("G:/My Drive/Wadden fish sounds/Data analysis/Final logs/LAUW_08_allsites_inputmod.xlsx"))
#logcompare<-as.data.frame(readxl::read_excel("G:/Shared drives/WaddenSea fish sounds/final log data janneke/R data and files wadden sea (Feb, May and June)/may1 for R.xlsx"))

#change names
soundnames<-unique(logfile$Comments)
soundnames
logfile$soundcode<-logfile$Comments
logfile$soundcode[logfile$soundcode=="hollow drumming" | logfile$soundcode=="knocking"]<-"Dr"
logfile$soundcode[logfile$soundcode=="bubble growl" | logfile$soundcode=="bubbel growl" | logfile$soundcode=="bubble growl\n" | logfile$soundcode=="buble growl\n" | logfile$soundcode=="bubbe growl"]<-"Sea-Gr"
logfile$soundcode[logfile$soundcode=="drum vibrations" | logfile$soundcode=="automatic vibrations" | logfile$soundcode=="woodpecker sound" | logfile$soundcode=="bowl growl" | logfile$soundcode=="low growl" | logfile$soundcode=="woodpecker"]<-"Grw"
logfile$soundcode[logfile$soundcode=="wavy frog" | logfile$soundcode=="low drumming" | logfile$soundcode=="fish croaking vertical"]<-"Cro"
logfile$soundcode[logfile$soundcode=="fish clicking"|logfile$soundcode=="lower fish clicking" | logfile$soundcode=="clicks" | logfile$soundcode=="fish clicking fade"]<-"Frt"
logfile$soundcode[logfile$soundcode=="fish grunt 4"]<-"GrBz"
logfile$soundcode[logfile$soundcode=="striped fish grunt"]<-"GrBz"
logfile$soundcode[logfile$soundcode=="water croaking"]<-"Bu-Frt"
logfile$soundcode[logfile$soundcode=="coned fish grunt" | logfile$soundcode=="solid fish grunt"]<-"Gr"
logfile$soundcode[logfile$soundcode=="frog beats"]<-"2Dr"
logfile$soundcode[logfile$soundcode=="reverse cod"]<-"higher fish grunt"
logfile$soundcode[logfile$soundcode=="bonking"]<-"LDr"
logfile$soundcode[logfile$soundcode=="double drum"]<-"2Dr"
logfile$soundcode[logfile$soundcode=="baby seal" | logfile$soundcode=="singular seal sound"]<-"BCa"
logfile$soundcode[logfile$soundcode=="chirping"]<-"mGr"

#remove noise period annotations
logfile2<-subset(logfile,soundcode!="very loud period")

require(ggplot2)
ggplot(data=logfile2,aes(x=soundcode))+geom_bar()
```
Include deployment duration and effort
```{r}
logfile2$depDurS<-0
logfile2$depDurS[logfile2$Location=="LAUW1_OFF"]<-836626
logfile2$depDurS[logfile2$Location=="LAUW1_ON"]<-835886
logfile2$depDurS[logfile2$Location=="LAUW2_OFF"]<-835475
logfile2$depDurS[logfile2$Location=="LAUW2_ON"]<-842109
```

```{r}
write.csv(logfile2,file="G:/Shared drives/WaddenSea fish sounds/Lauwersoog logs for publication/LAUW_08_allsites_soundcodesaligned.csv")
```

Improving the data file
```{r}
setwd("G:/Shared drives/WaddenSea fish sounds/Lauwersoog logs for publication")

file<-list.files(path="G:/Shared drives/WaddenSea fish sounds/Lauwersoog logs for publication")
filetype<-"allsites_soundcodesaligned.csv"
file.list<-grepl(filetype,file,fixed=TRUE)
file.WC<-file[file.list]

LAUWall<-NULL
for(n in 1:length(file.WC)){
  dat<-read.csv(file.WC[n])
  #titel<-file.WC[n]
  #tel<-length(dat[,1])
  #dat$file<-rep(titel,tel)
  #loc<-strsplit(titel,"_","]",2)
  #dat$Site<-rep(loc[[1]][2],tel)
  #dat$Deployment<-rep(loc[[1]][3],tel)
  LAUWall<-rbind(LAUWall,dat)
}
```
#Changing date time from local to UTC
```{r}
LAUWall$StartTime2<-as.POSIXct(LAUWall$StartTime,tz="CET")
LAUWall$StartUTC<-format(LAUWall$StartTime2,tz="UTC")

LAUWall$EndTime2<-as.POSIXct(LAUWall$EndTime,tz="CET")
LAUWall$EndUTC<-format(LAUWall$EndTime2,tz="UTC")
```


```{r}
summary(LAUWall)
LAUWsub<-LAUWall[,c(6:8,17:21,23,25)]
summary(LAUWsub)
table(LAUWsub$Location,LAUWsub$Deployment)
table(LAUWsub$Deployment,LAUWsub$soundcode)
write.csv(LAUWsub,file="G:/Shared drives/WaddenSea fish sounds/Lauwersoog logs for publication/LAUW_2023_soundcodesaligned_UTC.csv")
```

```{r}
ggplot(data=subset(RESsub,soundcode=="Dr"),aes(x=Deployment))+geom_bar(aes(fill=soundcode))
```

#Cleaning up Maryann's soundcodes so they are easier to compare
```{r}
soundsMW2$soundcode[soundsMW2$soundcode=="Background" | soundsMW2$soundcode=="B" | soundsMW2$soundcode=="Bckgrnd" | soundsMW2$soundcode=="Bkgrnd"]<-"Background"
soundsMW2$soundcode[soundsMW2$soundcode=="Bca"]<- "BCa"

#Subsetting Maryann's dataset to only include sounds that are underwater, biological and not background
soundsMWsub<-subset(soundsMW2,soundcode=="Gr" | soundcode=="Gr/Bz" | soundcode=="GrBz" | soundcode=="Kn / GrBz")
soundsJdBsub<-subset(soundsJdB, soundcode=="Gr" | soundcode=="GrBz" | soundcode=="Grw")

require(ggplot2)
soundsMWsub$y<-1
soundsJdBsub$y<-0.9
ggplot(data=soundsMWsub,aes(x=startUTC,y=y))+geom_point()+geom_point(data=subset(soundsJdBsub),aes(x=StartTime,y=y),colour="red")+ylim(0.5,1.5)
  xlim(as.POSIXct(c("2023-05-08 00:00:00","2023-05-09 12:25:59"),format="%Y-%m-%d %H:%M:%S"))
```
